A plot is worth a thousand tests: assessing residual diagnostics with visual inference

### 2023-03-13 paper_comparison/paper_comparison/paper_comparison.tex
compile pdf

\title{Why is it that statistical tests for residuals are not widely
used? An explanation using visual inference.}

@@

Title (jokes: finding the pattern in residual tests) (why using
visulization is important in regression diagnostics? an explanation
using visual inference) (a picture is worth a thousand word) (a plot is
worth a thousand test: )

(title refine - rank: low)

@@

\[E = \frac{1}{2}\left(\boldsymbol{\mu}_z'(diag(\boldsymbol{R}_a\sigma^2))^{-1}\boldsymbol{\mu}_z\right)\]

\noindent where
\(\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z\)
is the expected values of residuals with \(\boldsymbol{Z}\),
\(\boldsymbol{R}_a = \boldsymbol{I}_n - \boldsymbol{H}_a\) is the
residual operator,
\(\boldsymbol{H}_a = \boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'\)
is the hat matrix. For the non-linearity model this can be rewritten as

\[E = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a))\right)\]

\noindent where \(\boldsymbol{V}\) is the covariance matrix of the error
term.

@@

\[E = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a))\right)\]
\noindent where \(diag(.)\) is the diagonal matrix constructed from the
diagonal elements of a matrix.

@@

To compute the effect size for each lineup we simulate ??

@@

For non-linearity patterns, the power curve of RESET test climbs
aggressively from 16\% to 100\% as \(log_e(E)\) increases from 0 to 2.5,
while power of other tests respond inactively to the change of effect,
showing that RESET test is much more sensitive to weak non-linear
structure. Meanwhile, no noticeable residual departures can be spotted
from the example residual plots.

>>>>

\title{A plot is worth a thousand tests: assessing residual diagnostics
with visual inference}

@@

@@

\[E = \frac{1}{2}\left(\boldsymbol{\mu}_z'(diag(\boldsymbol{R}_a\sigma^2))^{-1}\boldsymbol{\mu}_z\right)\]
\noindent where \(diag(.)\) is the diagonal matrix constructed from the
diagonal elements of a matrix,
\(\boldsymbol{R}_a = \boldsymbol{I}_n - \boldsymbol{H}_a\) is the
residual operator,
\(\boldsymbol{H}_a = \boldsymbol{X}_a(\boldsymbol{X}_a'\boldsymbol{X}_a)^{-1}\boldsymbol{X}_a'\)
is the hat matrix,
\(\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z\)
is the expected values of residuals with \(\boldsymbol{Z}\) be any
higher order terms of \(\boldsymbol{X}\) leave out by the regression
equation and \(\boldsymbol{\beta}_z\) be the corresponding coefficients,
and \(\sigma^2\boldsymbol{I}\) is the assumed covariance matrix of the
error term when \(H_0\) is true.

@@

\[E = \frac{1}{2}\left(log\frac{|diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')|}{|diag(\boldsymbol{R}_a)|} - n + tr(diag(\boldsymbol{R}_a\boldsymbol{V}\boldsymbol{R}_a')^{-1}diag(\boldsymbol{R}_a))\right)\]
\noindent where \(\boldsymbol{V}\) is the actual covariance matrix of
the error term.

@@

To compute the effect size for each lineup we simulate a sufficient
large number of samples from the same model, in each sample, the number
of observations \(n\) is fixed. We then compute the effect size for each
sample and take the average as the final value. This ensures lineups
constructed with the same experimental factors will share the same
effect size.

@@

For non-linearity patterns, the power curve of RESET test climbs
aggressively from 7\% to 84\% as \(log_e(E)\) increases from 0 to 2.5,
while power of other tests respond inactively to the change of effect,
showing that RESET test is much more sensitive to weak non-linear
structure. Meanwhile, no noticeable residual departures can be spotted
from the example residual plots.
